Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-01-31 paper/1_Initial/manuscript.qmd
changes to manuscript

Various measures of interoception have been developed (see **@fig-measures**), forming a combination of "objective" and "subjective" assessments (i.e., physiological tasks such as the heart beat counting or tracking vs. questionnaires and subjective scales involving a metacognitive reflection), "explicit" and "implicit" paradigms (i.e., directing participants' awareness and attention to interoceptive processes *vs.* measuring interoception unbeknownst to them), various interoceptive modalities (e.g., cardioception, respiroception, gastroception) and theoretical dimensions (e.g., accuracy, sensitivity, awareness).
While there is no consensus as to which particular approach provides the most accurate and "pure" measure of interoception and interoceptive abilities (assuming it is a unidimensional construct), it is instead plausible that each measure has strengths and limitations, and a utility dependent on the context and goal at hand [@jahedi2014].

```{r}
#| label: "fig-measures"
#| fig-cap: "Different ways in which interoception can be measured."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

@@

The use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical. However, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information. These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively. This approach offers useful and interesting measures [@murphy2019; @lin2023].

The notion that self-reports might not reflect the same processes as other interoception measures is important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures].

<!-- lack of convergence between self-reported questionnaires -->

A recent systematic review by @desmedt2022measures examined whether various questionnaires designed to measure "interoceptive sensibility" truly assess the same construct. Notably, @garfinkel2015knowing defines interoceptive sensibility as the self-reported tendency to focus on and detect internal sensations, whereas @khalsa2018interoception defines it more narrowly, excluding the ability to detect these signals. According to @desmedt2022measures, most authors adopt the latter definition. The review found that these questionnaires measure related but distinct constructs, leading researchers to commit the 'jingle fallacy'—mistakenly treating them as equivalent measures of interoceptive sensibility.

**TODO: do a better job**

<!-- Moreover, the current subjective measures do not correlate well with the objective measures [@murphy2019]. -->

Few studies have examined the correlation between objective interoceptive measures, such as the Heartbeat Counting Task (HCT; Schandry, 1981) and the Heartbeat Detection Task (HDT; Kleckner et al., 2015). Existing findings report weak correlations for the HCT (Murphy et al., 2019), no correlations except for the sensitivity variable of the HCT (Brand et al., 2023), and small correlations for the HDT (Brand et al., 2023). These results suggest that the IAS, a subjective measure, does not strongly align with objective interoceptive assessments.

A better understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor structure for the IAS, the authors underline its acceptable but imperfect fit [@murphy2019; p. 127], calling on further investigation of the scale's factor structure. Notably, the authors did not define these factors, as no clear explanations were evident. They suggested that the first factor reflects the perception of interoceptive signals, while the second pertains to signals that may be difficult to perceive solely through interoceptive information.

<!-- [@morin2016] is a methodological paper on ESEM -->

@@

Discussions have also been focused on specific items. For instance, @murphy2019 notes that some items might measure direct interoceptive signals such as cardioception, while others might capture phenomena not perceivable through interoceptive signals alone (e.g., "bruising"; p. 119). 
@lin2023 also highlights their correlation analysis, showing five locally dependent pairs and three items (touch, blood sugar, bruise) with exceptionally high difficulty and low discrimination. 
Additionally, @campos2021 reported "tickle" to be the only item that reflected more specific factors than the general factor. 
Interestingly, @lin2023 reported that all items of the IAS grouped together using a new approach, Exploratory Graph Analysis [EGA; @golino2017exploratory], to assess convergent and discriminant validity, providing further evidence for unidimensionality.

<!-- Relationship with other interoception measures --> 
<!-- The MAIA was introduced above -->

The IAS has naturally been compared to other interoception-related scales, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional], except for the Non-Distracting and Not-Worrying subscales [@brand2023]. Interestingly, findings on the correlation between the IAS and the body awareness dimension of the Body Perception Questionnaire [BPQ-A, @porges1993body] have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019]. Small positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023]. Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [@lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia; @brand2023; @murphy2019]. Lastly, small positive correlations have also been reported with the Interoceptive Attention Scale [IATS, @lin2023; @koike2023], though studies have also found no correlation between these measures [@gabriele2022dissociations].

@@

While assessing the validity of an interoception scale can be conceived as theoretically challenging, several measures have been used to assess convergent validity for the the IAS, including expected negative associations with alexithymia [@brand2023; @koike2023; @campos2021, @murphy2019; @lin2023], somatic symptoms [@brand2023; @koike2023, @lin2023], depressive symptoms [@brand2023; @koike2023; @lin2023], anxiety [@brand2023], neuroticism [@brand2023] and self-esteem [@murphy2019]. 
<!-- **anything else?**. TO BE COMPLETED -- to be determined -->

@@

The current study aims at 1) clarifying the structure of the IAS with a meta-analytic approach that leverages existing data and contrast the traditional CFA/SEM factor-based analyses with network-based ones such as EGA.
2) The second part will provide an overview of the dispositional correlates of the IAS, providing an overview of the pattern of associations that is key to better understand the nature, place and role of interoception questionnaires within a larger context.

@@

Our search focused on studies citing the original IAS validation paper [@murphy2019], identifying 136 papers (as of 01/05/2024).
To qualify for inclusion, papers needed to (1) provide accessible data in open-access, (2) employ the IAS as a measure, and (3) report individual IAS items scores.
A total of 10 studies was included (see **Table 1**).
We also included the data of two unpublished (but already open-access) studies from the authors and one from another author.
The total N participants was 32,214 participants (*Mean* = 48.6 years, *SD* = 13.1, 71.6% Female).

<!-- \begin{landscape} -->
<!-- \input{../../analysis/figures/table1.tex} -->
<!-- \end{landscape} -->

@@

To examine the factor structure of the IAS, a two-step approach was employed.
First, Exploratory Graph Analysis (EGA), was used to estimate the dimensions via network estimation and community detection, alongside assessing the stability of dimensions and items using the bootstrapping techniques [@golino2017exploratory].
The selection of EGA was motivated by its capability to handle complex, multidimensional data and provide robust dimension estimates.
A novel network psychometrics - Unique variable analysis [UVA; @christensen2023unique] - approach based on the weighted topological overlap will be computed to evaluate which items have substantial local dependence (> 0.25).
Subsequently, exploratory factor analysis (EFA) was employed followed by confirmatory factor analysis (CFA).

@@

Visualizing the distribution of the items for all samples suggests the presence of a consistent modal value (@fig-distributions).
In other words, participants are most likely to answer 4/5 (i.e., agree) on most items (but "affective touch", "blood sugar", and "bruise" that exhibit a different distributional pattern).
Additionally, one can note the low density on extreme values (1 and 5), meaning that the bulk of answers (i.e., 99\%) varies between 3 values.
The interindividual variability seems improved in the samples using an analogue scale, displaying a more continuous and progressive spread of answers.

@@

When accounting for all samples, the factor analysis reveals that a 4-factor structure fits best.
The exploratory factor analysis revealed that 4 latent factors (oblimin rotation) accounted for 51.95% of the total variance of the original data (MR2 = 19.08%, MR1 = 17.78%, MR4 = 9.34%, MR3 = 5.75%).
Since UVA identified "tickle" as the item to be removed—and it also had the lowest uniqueness value in factor analysis with a similar loading to "itch"—"tickle" was excluded from subsequent analyses.

@@

Confirmatory factor analysis compared 5 models: a single-factor solution, a 4-factor solution, a 5-factor solution, a 6-factor solution and a 7-factor solution. The latter was preferred in most datasets, including with indices that penalise increased number of parameters (such as BIC). There was no evidence for higher order factors.

@@

These findings contrast with previous research, which all found that 2-factor model[@murphy2019; @koike2023], 1-factor model [@brand2023] and bifactor model [@lin2023; @campos2021] fits the data best.
While this analysis also revealed an okay fit for the 1-factor model, the 4-factor model revealed the best fit.
The 4-factor structure reveals different ‘hubs' of items that are related, not only in this structure analysis, but also in underlying mechanisms.
The ‘wind-burp-cough-sneeze-vomit' category, for example, only entails items that are linked to excretion through the mouth.
The other categories are organized similarly.
This organization and structure is useful for further analysis, as the data can be analyzed and interpreted according to a grouping that is coherent in result, as well as underlying mechanisms.

@@

There are several items that show redundancy suggesting that adapting the IAS would be beneficial for validity.
Based on the given results, we suggest removing the tickle, while keeping the itch item [**todo**: stats?].
Other items with slight redundancy were Hungry and Thirsty, Urinate and Defecate, and Sneeze and Cough.

@@

Interestingly, @lin2023 also found that tickle and itch were redundant, excluding one of them but the reason being that the character for both words is the same in the Chinese language.
On top of that, they came up with a shortened version of the IAS, excluding further items, resulting in a 12-item IAS, which aligns with our findings, suggesting that further items are ambiguous as to whether they should be removed.
In contrast, other findings also found itch and tickle to be redundant but did not suggest excluding items [@campos2021].

@@

The findings indicate a high proportion of answers at 4 (see Figure 1), especially when using a 5-step scale.
The analogue scale shows a more dispersed distribution, with some answers indicating the highest 5/5, which was not the case in the 5-step scale.
Therefore, we recommend using an analog scale for the IAS.

Before this paper, the IAS has not yet been used or analyzed with an analog scale, rather than a five step scale.
Therefore, this study provides a novel approach to improving the IAS in a simple manner.

@@

There are several limitations to the IAS; There are some redundant items, the 5-point scale does not provide great variability, and the structure could be improved.
Therefore, improving the IAS, or creating a new questionnaire investigating interoception could be useful to achieving reliable and accurate indication of interoceptive awareness.

@@

Study 2 aims to investigate correlates of the IAS.
Correlations of the IAS will be computed to assess the relationship between subjective interoceptive accuracy and other subjective measures of interoception, mood, psychopathology, personality, and beliefs.
Investigating correlates will help validate the IAS, as well as other interoceptive measures in the future.
The

@@

@@

<!-- \input{../../analysis/figures/table2.tex} -->

@@

Correlations will be computed using the correlation package under a Bayesian framework (ref).

### Results
**todo:  compute average correlations **

The EGA components capture groupings of pairs of items, such as "wind" and "burp", "cough" and "sneeze", or "muscle" and "pain".
These groupings were used in the correlational analysis, to observe how much each group/pair correlates with other factors, such as Alexithymia, or the MAIA (see Figure 2).

Alexithymia is negatively correlated with all interoceptive groups.
Autistic traits are mostly negatively correlated with IAS measure, except for patterns and numbers (as an autistic trait), which is significantly and positively correlated with the "itch/bruise" pairing.
The BPQ Body Awareness part is positively correlated with all IAS pairs, whereas the Autonomic Reactivity part is negatively correlated with all IAS groupings.
Conspiracy Beliefs were all positively correlated with the IAS pairs, however, only Global Conspiracy with "hungry/thirsty", "urinate/defecate" and "cough/sneeze", as well as personal wellbeing with "hungry/thirsty", and Information Control with "cough/sneeze" were significantly positively correlated.
Demographic data is also mostly positively correlated with the IAS findings, where gender and age more strongly correlated with "hungry/thirsty".
In this analysis, lying profile is not strongly correlated with the IAS; Except for contextuality, which shows a significant negative correlation with "itch/bruise".
The MAIA has a strong positive correlation with most IAS pairings, except for the "not worrying" and "not distracting" items of the MAIA, which show less strong, or even negative correlations with all IAS item pairings.
Maladaptiveness had mostly negative correlations with the IAS, with only a few significant correlations, namely "psychoticism", "negative affect", and "detachment" with "muscle/pain" and "hungry/thirsty", as well as "negative affect" with "wind/burp".
Overall, mood was mostly negatively correlated with the IAS, where "hungry/thirsty" had the strongest negative correlation of all mood measures.
Except for "neuroticism" and "honesty-humility", personality traits, such as "openness" and "extraversion" were positively correlated with the IAS groupings.
Shizotypic traits were mostly negatively correlated with the IAS, with "hungry/thirsty" showing the strongest negative correlation between schizotypic and the IAS.
World beliefs were mostly positively correlated with the IAS, however, only a few were significant: "hierarchical", "enticing", and "alive" correlates significantly with "muscle/pain"; "Understandable" has a significant positive correlation with "heart/breathing"; And "hierarchical" and "alive" has a significant positive correlation with "hungry/thirsty".

@@

Our findings confirmed that interception lies within an intricate network of correlates.
Alexithyma has the most negative correlation with the IAS, whereas, the MAIA questionnaire is most strongly positively correlated with the IAS.
The correlates cannot only explain attributes of interoception, but can also be used to validate measures for interoception.

While these results reveal correlations of different categories with the IAS, the findings are limited to the given questionnaire.
However, they give a good indication of how interoception in general might be linked to the different categories, leading to interesting findings.
The results show a consistent pattern of correlations with other interoception measures, psychopathology, and highlighting of interesting exploratory results, such as how primal world beliefs correlate with the IAS.

The analysis revealed a negative correlation between alexithymia and scores in the IAS, which is in line with previous research [@herbert2011].
A negative correlation between autism and interoceptive awareness, which has been found in this sample, has also been found in previous research [@dubois2016].
Conspiracy Beliefs and IAS scores were not strongly correlated.
However, there was a slight positive correlation.
To the best of our knowledge, this interaction has not been investigated before.
However, research shows correlations between (political) beliefs and interoception, which could underlie the same mechanism [@ruisch2022].
Lying profile was also not strongly correlated with the IAS.
Previous research provides some evidence for a relationshipbetween interoception and lying profile [@makowski2023a], indicating a contrast from the literature to the results at hand.
The analysis showed that mood and the IAS scores have a strong negative correlation, which has been investigated in previous studies, finding a correlation between the two variables as well [@solanolópez2018].
The analysis showed that personality is also correlated with interoceptive accuracy scores as retrieved by the IAS.
While there are many aspects of personality, put simply, personality and interoception have been found to correlate in previous studies as well [@erle2021].
Our findings show some negative correlations between schizotypy and interoception.
Previous research has investigated this relationship and, similarly, found a negative correlation between interoceptive awareness and schizotypy, particularly at the onset of psychosis [@torregrossa2022].
Lastly, World Beliefs and Interoception showed some significant positive correlations.
World Beliefs have not been linked with interoception previously.
As mentioned, other beliefs, such as political beliefs have been mentioned to interoception, where significant correlations were observed [@ruisch2022a].
Further research is needed to understand this mechanism and validate whether world beliefs, which lay the framework through which we see the world [@clifton2020], is needed.

The above findings show how interoception is correlated with many lifestyle and health factors, and is therefore an important concept to study.
Here we showed many correlates of the IAS and lifestyle, as well as health factors.
This is not only valuable to understand interception, and the role it plays in our lives, better, but also crucial to further validate the Ias, as well as other interoception measures.
This analysis therefore lays the groundwork for validating new interoceptive measures and questionnaires needed to grasp interoception and its role in our life.

@@

The analyses revealed that the IAS has 4-factor structure, and a rather uneven distribution.
The findings indicate that the IAS is measuring interoception in an acceptable manner, but there is also room for improvement.
Furthermore, the results reveal different correlation measures with the IAS, giving space to further explore whether and how measures correlate with interoception/the IAS.
In the following section, the IAS is discussed further to reveal shortcomings and strengths of the questionnaire.
Finally, future steps will be revealed to improve the way in which we measure interoception.

Overall, the IAS is straightforward with its sensation-centered items.
There are several points of improvement to the IAS, which this work suggests.
Firstly, removing redundant items, such as the itch item as shown in the above analysis.
Previous work has also shown that the itch and tickle items are subject to redundancy [@lin2023, @campos2021] has shown.
Interestingly, @campos2021 suggest removing the tickle item instead.
@lin2023 on the other hand removed the itch item, because the sign for itch and tickle is the same.
Furthermore, this work recommends using analog scales, instead of 5 point scales.
In a 5-point scale, the variability is clearly limited with most people choosing ⅘.
As visible in Figure X, the variability increases significantly, using an analog scale.
However, it is important to note that even with an analog scale, the variability of the IAS is still limited.
The more variability a scale shows, the more results can depict individual differences between participants.
Therefore, good dispersion is essential for achieving relevant and usable results, and striving for more variability would be beneficial to this interoception scale, too.

>>>>

Various measures of interoception have been developed (see @fig-measures), forming a combination of "objective" and "subjective" assessments (i.e., physiological tasks such as the heart beat counting or tracking vs. questionnaires and subjective scales involving a metacognitive reflection), "explicit" and "implicit" paradigms (i.e., directing participants' awareness and attention to interoceptive processes *vs.* measuring interoception unbeknownst to them), various interoceptive modalities (e.g., cardioception, respiroception, gastroception) and theoretical dimensions (e.g., accuracy, sensitivity, awareness).
While there is no consensus as to which particular approach provides the most accurate and "pure" measure of interoception and interoceptive abilities (assuming it is a unidimensional construct), it is instead plausible that each measure has strengths and limitations, and a utility dependent on the context and goal at hand [@jahedi2014].

```{r}
#| warning: false 
#| label: "fig-measures"
#| fig-cap: "Different ways in which interoception can be measured."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
## | out-width: "100%"

@@

The use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical.
However, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information.
These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively [@khalsa2018interoception; @suksasilp2022towards].
This approach offers useful and interesting measures [@murphy2019; @lin2023].

<!-- lack of correlation between objective and subjective measures + lack of correlation between self-report measures-->

The notion that self-reports might not reflect the same processes as other interoception measures is important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures].
While few studies have explored the correlation between objective interoceptive measures,  such as the Heartbeat Counting Task [HCT, @schandry1981heart] and the Heartbeat Detection Task [HDT, @kleckner2015methodological], and correspondent self-report measures, existing findings typically show weak or no correlations [e.g., task-based accuracy vs. self-reported accuracy, @murphy2019; @brand2023].

Even among self-report measures low correlations have been observed, suggesting that these measures, while group in the same umbrella term of self-report assess different facets of interoception. 
A recent systematic review by @desmedt2022measures examined whether various questionnaires designed to assess "interoceptive sensibility" truly measure the same construct. 
Notably, @garfinkel2015knowing defines interoceptive sensibility as the self-reported tendency to focus on and detect internal sensations, whereas @khalsa2018interoception defines it more narrowly, excluding the ability to detect these signals. According to @desmedt2022measures, most authors adopt the latter definition. 
The review found that these questionnaires measure related but distinct constructs, leading researchers to mistakenly treating them as equivalent measures of interoceptive sensibility. 
A better understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor structure for the IAS, the authors underline its acceptable but imperfect fit [@murphy2019, p. 127], calling on further investigation of the scale's factor structure.
Notably, the authors did not define these factors, as no clear explanations were evident.
They suggested that the first factor reflects the perception of interoceptive signals, while the second pertains to signals that may be difficult to perceive solely through interoceptive information.

**note to dom**:@morin2016 is a methodological paper on ESEM

@@

Discussions have also been focused on specific items.
For instance, @murphy2019 notes that some items might measure direct interoceptive signals such as cardioception, while others might capture phenomena not perceivable through interoceptive signals alone (e.g., "bruising"; p. 119).
@lin2023 also highlights their correlation analysis, showing five locally dependent pairs and three items (touch, blood sugar, bruise) with exceptionally high difficulty and low discrimination.
Additionally, @campos2021 reported "tickle" to be the only item that reflected more specific factors than the general factor.
Interestingly, @lin2023 reported that all items of the IAS grouped together using a new approach, Exploratory Graph Analysis [EGA, @golino2017exploratory], to assess convergent and discriminant validity, providing further evidence for unidimensionality.

<!-- Relationship with other interoception measures -->

The IAS has naturally been compared to other interoception-related scales, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional], except for the Non-Distracting and Not-Worrying subscales [@brand2023].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the Body Perception Questionnaire [BPQ-A, @porges1993body] have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019].
Small positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [@lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia; @brand2023; @murphy2019].
Lastly, small positive correlations have also been reported with the Interoceptive Attention Scale [IATS, @lin2023; @koike2023], though studies have also found no correlation between these measures [@gabriele2022dissociations].
Interestingly, the IAS and the IATS supposedly measure two different interoceptive processes (i.e., accuracy and attention, respectively) which contradict @murphy2019 proposed 2x2 framework.

@@

While assessing the validity of an interoception scale can be conceived as theoretically challenging, several measures have been used to assess convergent validity for the the IAS, including expected negative associations with alexithymia [@brand2023; @koike2023; @campos2021; @murphy2019; @lin2023], somatic symptoms [@brand2023; @koike2023; @lin2023], depressive symptoms [@brand2023; @koike2023; @lin2023], anxiety [@brand2023], neuroticism [@brand2023] and self-esteem [@murphy2019].
<!-- **anything else?**. TO BE COMPLETED -- to be determined -->

@@

The current study aims at 1) clarifying the structure of the IAS with a meta-analytic approach that leverages existing data and contrast the traditional CFA/SEM factor-based analyses with network-based ones such as EGA;
2) the second part will provide an overview of the dispositional correlates of the IAS, clarifying the pattern of associations which is key to better understand the nature, place and role of interoception questionnaires within a larger context.

@@

Our search focused on studies citing the original IAS validation paper [@murphy2019], identifying 136 papers (as of 01/05/2024).
To qualify for inclusion, papers needed to (1) provide accessible data in open-access, (2) employ the IAS as a measure, and (3) report individual IAS items scores.
A total of 10 studies was included (see **Table 1**).
We also included the data of two unpublished (but already open-access) studies from the authors and one from another researcher.
The total N participants was 32,214 participants (*Mean* = 48.6 years, *SD* = 13.1, 71.6% Female).

\begin{landscape}

\input{../../analysis/figures/table1.tex}

\end{landscape}

@@

To examine the factor structure of the IAS, a two-step approach was employed.
First, Exploratory Graph Analysis (EGA), was used to estimate the dimensions via network estimation and community detection, alongside assessing the stability of dimensions and items using the bootstrapping techniques [@golino2017exploratory].
The selection of EGA was motivated by its capability to handle complex, multidimensional data and provide robust dimension estimates.
A novel network psychometrics - Unique variable analysis [UVA, @christensen2023unique] - approach based on the weighted topological overlap will be computed to evaluate which items have substantial local dependence (\> 0.25).
Subsequently, exploratory factor analysis (EFA) was employed followed by confirmatory factor analysis (CFA).

@@

Visualizing the distribution of the items for all samples suggests the presence of a consistent modal value (@fig-distributions).
In other words, participants are most likely to answer 4/5 (i.e., agree) on most items (but "affective touch", "blood sugar", and "bruise" that exhibit a different distributional pattern).
Additionally, one can note the low density on extreme values (1 and 5), meaning that the bulk of answers (i.e., 99%) varies between 3 values.
The interindividual variability seems improved in the samples using an analogue scale, displaying a more continuous and progressive spread of answers.

@@

<!-- check EFA results again -->

When accounting for all samples, the factor analysis reveals that a 4-factor structure fits best.
The exploratory factor analysis revealed that 4 latent factors (oblimin rotation) accounted for  41.58% of the total variance of the original data (MR1 = 14.45%, MR3 = 11.76%, MR2 = 8.09%, MR4 = 7.28%).
Since UVA identified "tickle" as the item to be removed—and it also had the lowest uniqueness value in factor analysis with a similar loading to "itch"—" it was excluded from subsequent analyses.

@@

CFA was computed with the removal of the item "tickle" as it was constantly flagged as redundant. This analysis compared 5 models: a single-factor solution, a 4-factor solution, a 5-factor solution, a 6-factor solution and a 7-factor solution. 
The latter was preferred in most datasets, including with indices that penalise increased number of parameters (such as BIC).
There was no evidence for higher order factors.

@@

These findings contrast with previous research, which all found that 2-factor model [@murphy2019; @koike2023], 1-factor model [@brand2023] and bifactor model [@lin2023; @campos2021] fits the data best.
While this analysis also revealed an okay fit for the 1-factor model, the 4-factor model was superior.
The 4-factor structure reveals different ‘hubs' of items that are related, not only in this structure analysis, but also in underlying mechanisms.
The ‘wind-burp-cough-sneeze-vomit' category, for example, only entails items that are linked to excretion through the mouth.
The other categories are organized similarly.
This organization and structure is useful for further analysis, as the data can be analyzed and interpreted according to a grouping that is coherent in result, as well as underlying mechanisms.

@@

**note to dom**: what stats do you mean here? the UVA one?

There are several items that show redundancy suggesting that adapting the IAS would be beneficial for validity.
Based on the given results, we suggest removing the tickle, while keeping the itch item  \[**todo**: stats?\].
Other items with slight redundancy were "hungry" and "thirsty", "urinate" and "defecate", and "sneeze" and "cough".

@@

Interestingly, @lin2023 also found that "tickle" and "itch" were redundant, which led them to excludw one of them. Although, the reason being that the character for both words is the same in the Chinese language.
On top of that, they came up with a shortened version of the IAS, excluding further items, resulting in a 12-item IAS, which aligns with our findings, suggesting that further items are ambiguous as to whether they should be removed.
Their 12-item IAS included the "hunger", "breath", "urinate", "taste", "vomit", "cough", "temperature", "sexual arousa", "wind", "muscle", "pain", "itch" items.  
In contrast, other findings also found "tickle" to be redundant but did not suggest excluding items [@campos2021].

@@

The findings indicate a high proportion of answers at 4 (see Figure 1), especially when using a 5-step scale.
The analogue scale shows a more dispersed distribution, with some answers indicating the highest 5/5, which was not the case in Likert-scales. 
Therefore, we recommend using an analog scale for the IAS.

Before this paper, the IAS had not yet been used or analyzed with an analog scale, rather than a five-step scale.
Therefore, this study provides a novel approach to improving the IAS in a simple manner.

@@

<!-- pooor arguments --> 
There are several limitations to the IAS. There are some redundant items, the 5-point scale does not provide great variability, and the structure could be improved.
Therefore, improving the IAS, or creating a new questionnaire investigating interoception could be useful to achieving reliable and accurate indication of interoceptive awareness.

@@

Study 2 aims to investigate correlates of the IAS.
Correlations of the IAS will be computed to assess the relationship between subjective interoceptive accuracy and other subjective measures of interoception, mood, psychopathology, personality, and beliefs.
Investigating correlates will help validate the IAS, as well as other interoceptive measures in the future.

@@

**note to dom** how do you want to present the questionnaires? in text format or in the table with their references on a footnote?

@@

\input{../../analysis/figures/table2.tex}

##### Interoceptive Measures

@@

Correlations will be computed using the correlation package under a Bayesian framework [@ben2020effectsize].

**todo: compute average correlations**

The EGA components capture groupings of pairs of items, such as "wind" and "burp", "cough" and "sneeze", and "muscle" and "pain".
These groupings were used in the correlational analysis, to observe how much each group/pair correlates with other factors, such as Alexithymia, or the MAIA (see Figure 2).

Alexithymia is negatively correlated with all interoceptive groups.
Autistic traits are mostly negatively correlated with IAS measure, except for patterns and numbers (as an autistic trait), which is significantly and positively correlated with the "itch/bruise" pairing.

The BPQ Body Awareness part is positively correlated with all IAS pairs, whereas the Autonomic Reactivity part is negatively correlated with all IAS groupings.
Conspiracy Beliefs were all positively correlated with the IAS pairs, however, only Global Conspiracy with "hungry/thirsty", "urinate/defecate" and "cough/sneeze", as well as personal wellbeing with "hungry/thirsty", and Information Control with "cough/sneeze" were significantly positively correlated.

Demographic data is also mostly positively correlated with the IAS findings, where gender and age more strongly correlated with "hungry/thirsty".

In this analysis, lying profile is not strongly correlated with the IAS, except for contextuality which shows a significant negative correlation with "itch/bruise".
The MAIA has a strong positive correlation with most IAS pairings, except for the Not Worrying and Not Distracting items of the MAIA, which show less strong, or even negative correlations with all IAS item pairings.

Maladaptiveness had mostly negative correlations with the IAS, with only a few significant correlations, namely "psychoticism", "negative affect", and "detachment" with "muscle/pain" and "hungry/thirsty", as well as "negative affect" with "wind/burp".

Overall, mood was mostly negatively correlated with the IAS, where "hungry/thirsty" had the strongest negative correlation of all mood measures.
Except for "neuroticism" and "honesty-humility", personality traits, such as "openness" and "extraversion" were positively correlated with the IAS groupings.
Shizotypic traits were mostly negatively correlated with the IAS, with "hungry/thirsty" showing the strongest negative correlation between schizotypic and the IAS.

World beliefs were mostly positively correlated with the IAS, however, only a few were significant: "hierarchical", "enticing", and "alive" correlates significantly with "muscle/pain"; "Understandable" has a significant positive correlation with "heart/breathing"; And "hierarchical" and "alive" has a significant positive correlation with "hungry/thirsty".

@@

<!-- strongest and weakest correlations with the IAS--> 
Our findings confirm that interoception exists within a complex network of correlates. 
Among these, alexithymia exhibits the strongest negative correlation with the IAS, whereas the MAIA questionnaire shows the strongest positive correlation. 
These correlates not only help explain different aspects of interoception but also serve as valuable tools for validating interoceptive measures.

<!-- consistent pattern of results and interesting exprolatory analysis --> 
While our results reveal various correlations with the IAS, they are limited to the scope of the given questionnaire.
Nonetheless, they provide valuable insights into how interoception may relate to different psychological and personality traits. 
The results show a consistent pattern of correlations with other measures and highlight interesting exploratory results, such as correlations between primal world beliefs with the IAS.

<!-- negative correlation with alexithymia and autism --> 
Our analysis found a strong negative correlation between alexithymia and IAS scores, aligning with previous research [@herbert2011; @brand2023; @murphy2019]. 
Similarly, a negative correlation between autism and interoceptive awareness was observed in our sample, consistent with prior findings [@dubois2016].

<!-- weak correlation with conspiracy beliefs --> 
Conspiracy beliefs did not strongly correlate with IAS scores, though a slight positive correlation was present. 
To our knowledge, this relationship has not been previously explored. 
However, prior studies have suggested connections between interoception and (political) beliefs, potentially pointing to shared underlying mechanisms [@ruisch2022].

<!-- weak correlation with lying profiles --> 
The relationship between interoception and lying profiles was also weak. 
This contrasts with previous research suggesting associations between interoception and deception [@makowski2023a], warranting further investigation.

<!-- stroong negative correlations with Mood --> 
Mood and IAS scores exhibited a strong negative correlation, consistent with prior studies that have documented similar findings [@solanolópez2018]. 
Additionally, personality traits correlated with interoceptive accuracy scores, reinforcing existing research linking personality and interoception [@erle2021].

<!-- negative correlations with schizotypy --> 
We also observed negative correlations between schizotypy and interoception, in line with previous studies that identified a similar relationship with interoceptive awareness, particularly in individuals at risk for psychosis [@torregrossa2022].

<!-- positive correlations with some Primal World Beliefs --> 
Interestingly, world beliefs demonstrated significant positive correlations with interoception. 
While this relationship has not been previously documented, other forms of belief, such as political ideology, have been linked to interoception [@ruisch2022a]. 
Further research is needed to determine whether world beliefs, which shape our perception of reality [@clifton2020], are meaningfully connected to interoception.

<!--Summary -->

Overall, our findings highlight the broad relevance of interoception across various cognitive and affective traits, underscoring its significance in both research and clinical contexts. 
By identifying numerous correlates of the IAS, we contribute not only to a deeper understanding of interoception’s role in daily life but also to the ongoing validation of the IAS and other interoceptive measures. 
This analysis lays an important foundation for the development of new interoceptive assessment tools, further advancing our comprehension of interoception and its impact on human experience.

@@

<!-- factor structure -->

Our analyses revealed that the IAS follows a four-factor structure with an uneven distribution. 
While the findings indicate that the IAS measures interoception adequately, there is room for improvement. 
Additionally, different correlation measures with the IAS suggest opportunities for further exploration of how interoception is assessed.
In the following section, we discuss the strengths and shortcomings of the IAS, followed by proposed steps to enhance interoception measurement.

<!-- redundancy of tickle and other items -->

Overall, the IAS is straightforward in its sensation-centered items.
However, several areas for improvement emerge from this study. Firstly, redundant items should be removed, such as the "itch" item, as highlighted in our analysis. 
Previous research also suggests redundancy between itch and tickle items [@lin2023, @campos2021]. 
Interestingly, while @campos2021 does not recommend the removal of either, @lin2023 argues for removing the itch item due to their overlapping character representation.

<!-- benefit of analog scales -->

Furthermore, this study recommends using analog scales instead of 5-point scales. 
The limited variability of the 5-point scale often results in most responses clustering around 3 or 4. 
As shown in @fig-distributions, adopting an analog scale significantly increases variability. However, even with an analog scale, IAS variability remains constrained. 
Greater variability allows for better differentiation among participants, making dispersion an essential factor for obtaining meaningful results. 
Enhancing variability would therefore be beneficial for the IAS.

<!-- Underrepresentation of modalities -->

Despite these improvements, certain limitations persist in the IAS that affect its accuracy. 
Notably, some modalities are underrepresented—for instance, heart perception is measured by only one item. 
Expanding modality coverage would enhance variability within each category, leading to more nuanced results. 
Moreover, the IAS lacks a clear theoretical or empirical structure, with only small item groupings. 
Ideally, a scale should allow for clear groupings that support meaningful data analysis. 
In this study, each group contained only two items, resulting in low scores and limited variability. 
Additionally, some IAS items are ambiguous, with their interpretation depending on context. 
For example, an item about perceiving heartbeats and another about vomiting could both relate to anxiety, leading to results that may differ from initial expectations. 
Thus, the grouping and structure of the IAS require refinement.

<!-- positve response bias from only positive worded items -->
